                             Project Assignment- 1
======================
### Mean total steps taken per day
Reading and having a first look at data
```{r}
activity <- read.csv("activity.csv")
names(activity)
str(activity)
head(activity)
```
Total steps taken in a day
```{r}
totalsteps <- aggregate(steps~date, data = activity, FUN = sum, na.rm=T)
# Have a look at dataset
head(totalsteps)
#plotting graph
hist(totalsteps$steps, xlab = "Steps per day", main = "total number of steps per day", col = "red")
```
Mean and median steps
```{r}
meansteps <- mean(totalsteps$steps)
mediansteps <- median(totalsteps$steps)
meansteps
mediansteps
```
### Average daily activity pattern

```{r}
intervalsteps <- aggregate(steps~interval, data = activity, FUN=mean, na.rm = T)
#have a look at dataset
str(intervalsteps)
head(intervalsteps)
plot(intervalsteps$interval, intervalsteps$steps, col = "green", xlab = "Intervals", ylab = "Total Steps per interval", main = "Number of steps per interval", type = "l")
```
Finding max. steps
```{r}
maxsteps <- max(intervalsteps$steps)
#Maximun number of steps aken in a 5 min. interval is:
maxsteps
maxinterval <- intervalsteps$interval[which(intervalsteps$steps == maxsteps)]
#interval in which max. steps are taken is:
maxinterval
```
### Imputing missing vaues
```{r}
#The number of missing values in out activity dataset is:
sum(is.na(activity))
```
Strategy for filling the NA's
```{r}
missingvalues <- subset(activity, is.na(steps))
par(mfrow = c(2,1), mar = c(3,3,2,2))
hist(missingvalues$interval, main = "NA's per interval")
hist(as.POSIXct(missingvalues$date), main = "NA's per date", breaks = 61)
#we can see that NA's run equally over all interval but in days it is available for only 8 days, so we will take of mean of date across all dataset
```

Creating new data set with missing values
```{r}
meanstepinterval <- tapply(activity$steps, activity$interval, mean, na.rm = T)
activityna <- activity[is.na(activity$steps),]
activitynonna <- activity[!is.na(activity$steps),]
activityna$steps <- as.factor(activityna$interval)
levels(activityna$steps) <- meanstepinterval
levels(activityna$steps) <- round(as.numeric(levels(activityna$steps)))
activityna$steps <- as.integer(as.vector(activityna$steps))
imputedactivity <- rbind(activityna, activitynonna)
str(imputedactivity)
sum(is.na(imputedactivity))
# Thus all the NA's have been removed
```

Histogram of total number of steps taken in each day

```{r}
par(mfrow = c(1,2))
activitysteps <- aggregate(steps~date, data = activity, FUN = sum, na.rm =T)
hist(activitysteps$steps, xlab = "Steps per day", main ="Total steps per day, no NA's", col = "blue")
impactivitysteps <- aggregate(steps~date, data = imputedactivity, FUN=sum)
hist(impactivitysteps$steps, xlab ="Steps per day", main ="NA's imputed, Total steps per day", col = "blue")

imputedmean <- mean(impactivitysteps$steps)
imputedmedian <- median(impactivitysteps$steps)
imputedmean
imputedmedian

result_mm <- data.frame(c(meansteps, mediansteps), c(imputedmean, imputedmedian))
colnames(result_mm) <- c("NA removed", "Imputed NA values")
rownames(result_mm) <- c("mean", "median")
result_mm

#Conclusion : imputing NA has no big effect on mean and median
```

###Difference in weekands and weekdays

```{r}
imputedactivity$daytype <- ifelse(weekdays(as.Date(imputedactivity$date)) == c("Saturday", "Sunday"), "Weekend", "weekday")
head(imputedactivity)
imputedactivity$daytype <- factor(imputedactivity$daytype)
stepsdaytype <- aggregate(steps~interval + daytype, data = imputedactivity, FUN = mean)
head(stepsdaytype)

library(ggplot2)
g <- ggplot(stepsdaytype, aes(interval, steps))
g + geom_line(col = "green") + facet_grid(daytype~.) + labs(x="Intervals", y = "Average Steps", title = "Activity Pattern")
```









### Conclusion : The subject gets an early start on weekdays then weekends





